---
title: "Writing BGM"
author: "Michael Sumner"
date: "24 November 2015"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## BGM from Spatial data in R

Create a dummy Spatial polygon layer (could be read from shapefile or whatever). 

```{r}
library(graticule)

## polygon map in LAEA

pg <- graticule(seq(50, 110, length = 8), seq(-70, -40, length = 6), proj = "+proj=laea +ellps=WGS84 +lon_0=65 +lat_0=-90", tiles = TRUE)
```

Populate with suitable or dummy values for Atlantis. 

```{r}
mkcent <- function(g) {
 x <- as.data.frame(coordinates(rgeos::gPointOnSurface(g, byid = TRUE)))
 names(x) <- c("insideX", "insideY")
 x
}
cn <- mkcent(pg)
pg[[names(cn)[1]]] <- cn[[1]]
pg[[names(cn)[2]]] <- cn[[2]]


```

Let's say that the outside margin of polygons are boundary boxes, excluding the bottom "row". 

```{r}

library(gris)
gr <- gris(pg)

## find outer verts (only those branches where a vertex occurs just once)
# out <- (gr$bXv %>% inner_join((gr$v  %>% inner_join(gr$bXv)  %>% group_by(.vx0)  %>% summarize(nv  = n())  %>% filter(nv == 1))) %>% 
#   select(.br0) %>% inner_join(gr$b) %>% distinct(.br0))$.ob0
# 
# plot(gr[out, ])
# 
# gr$o$boundary <- 0
# gr$o$boundary[out] <- 1
plot(gr)
text(coordinates(pg), lab = seq(nrow(gr$o)))
gr$o$boundary <- 0
out <- c(1:7, 8, 14, 15, 21, 22, 28, 29, 35)
gr$o$boundary[out] <- 1
## pull out boxes
allbox <- gr$o
## first must be a boundary
isbound <- allbox$boundary == 1
ord <- c(which(isbound)[1], which(!isbound), which(isbound)[-1])
allbox <- allbox[ord, ]
allbox$o$box <- seq(nrow(allbox$o)) - 1
#allverts <- allbox %>% inner_join(gr$b, c(".ob0" = ".ob0")) %>% inner_join(gr$bXv, c(".br0" =".br0")) %>% inner_join(gr$v, c(".vx0" = ".vx0")) 
mkbox <- function(x, num) {
  enames <- c("label", "inside", "nconn", "iface", "ibox", "botz", "area", "vertmix", "horizmix")
  # Box number 0
# box0.label	Box0
# box0.inside	96.61082843655237 -58.71767790250464
# box0.nconn	8
# box0.iface	15 29 28 26 25 23 22 6 
# box0.ibox	21 21 22 23 23 24 24 24 
# box0.botz	0.0
# box0.area	144.94569776898845
# box0.vertmix	0.0
# box0.horizmix	0.0
# box0.vert 111.3499999997004 -62.24999999960016
# box0.vert
  
  sprintf("box%i.", "", "") 
}
for (i in seq(nrow(allbox))) {
    
}
